---
title: "Chapter 6 - Working with CAS Tables"
author: "Yue Qi"
output: github_document
---

```{r setup}
knitr::opts_chunk$set()
```



Import the SWAT package and get connected.

```{r}
library("swat")
library(dplyr)
conn <- CAS('rdcgrdc.unx.sas.com', 39935)
cas.sessionProp.setSessOpt(conn, caslib = "casuser")
```

# Using CASTable Objects like a Data Frame

## CAS Table Introspection
check the class of iris data to make sure it is a data.frame instance
```{r}
class(iris)
```

Get the column names
```{r}
names(iris)
```

Load iris data.frame into a CAS table
```{r}
tbl <- as.casTable(conn, iris)
```

Show the class of tbl object
```{r}
class(tbl)
names(tbl)
```

```{r}
nrow(tbl)
dim(tbl)
```

```{r}
attributes(tbl)
```

```{r}
head(tbl, n = 3L)
```

```{r}
tail(tbl, n = 3L)
```

## Computing Simple Statistics
Run the summary function on a data.frame
```{r}
summ <- summary(iris)
summ
```

```{r}
class(summ)
```

Run the summary function on a CASTable
```{r}
cassumm <- summary(tbl)
cassumm
```

```{r}
class(cassumm)
```

```{r }
cas.sessionProp.setSessOpt(conn,caslib="HPS")
out  <- cas.table.loadTable(conn, path = 'megacorp5m.sashdat',casOut=list(replace = TRUE) )
cas.table.tableInfo(conn, 'MEGACORP5M')
```

```{r }
mega = defCasTable(conn,'MEGACORP5M')
system.time({ result = summary(mega) })
```

```{r }
result
```

```{r}
cas.sessionProp.setSessOpt(conn,caslib="casuser")
cas.count(tbl)
```

```{r}
cas.mean(tbl)
```

```{r}
cas.var(tbl)
```

```{r}
plot(tbl$Sepal.Length, tbl$Sepal.Width, pch = 16, type = 'p', 
  col = 'blue', main = "Plot on a CASTable")

```

```{r}
par(mfrow=c(1,3))

tbl_tmp <- tbl[tbl$Species == 'setosa',]
plot(tbl_tmp$Sepal.Length, tbl_tmp$Sepal.Width, pch = 16, type = 'p', 
    col = 'red', main = "setosa")

tbl_tmp <- tbl[tbl$Species == 'versicolor',]
plot(tbl_tmp$Sepal.Length, tbl_tmp$Sepal.Width, pch = 17, type = 'p', 
    col = 'green', main = "versicolor")

tbl_tmp <- tbl[tbl$Species == 'virginica',]
plot(tbl_tmp$Sepal.Length, tbl_tmp$Sepal.Width, pch = 18, type = 'p', 
    col = 'blue', main = "virginica")

```

# Sorting, Data Selection, and Iteration
## Fetching Data with a Sort Order
```{r}
sorttbl <- defCasTable(conn, 'iris', orderby = 'Sepal.Length')
```

```{r}
sorttbl <- defCasTable(conn, 'iris')
sorttbl@orderby <- 'Sepal.Length'

```

```{r}
head(sorttbl, n = 10L)
```

```{r}
tail(sorttbl, n = 10L)
```
## Iterating through Columns
```{r}
for (col in names(sorttbl)){
  print(col)
}

```
## Techniques for Indexing and Selecting Data
### Selecting Columns by Label and Position
```{r}
sorttbl@orderby <- 'Sepal.Width'
col <- sorttbl['Sepal.Width']
head(col)

```

```{r}
col <- sorttbl$Sepal.Width
head(col)

```

```{r}
widths <- sorttbl[c('Sepal.Width', 'Petal.Width', 'Species')]
class(widths)

```

```{r}
head(widths)
```

```{r}
summary(widths)
```

```{r}
cas.table.columnInfo(widths)
```

```{r}
sorttbl@orderby <- 'Sepal.Width'
```
Select the petal_width column
```{r}
head(sorttbl[,"Petal.Width"])

```

Select a list of columns
```{r}

head(sorttbl[,c("Petal.Width","Sepal.Length")])

```

```{r}
head(sorttbl[,3])
```

```{r}
head(sorttbl[,1:3])
```

```{r}
head(sorttbl[,c(4,2)])
```
### Dynamic Data Selection
```{r}
expr <- sorttbl$Petal.Length > 6.5
```

```{r}
newtbl <- sorttbl[expr,]
head(newtbl)

```

```{r}
newtbl <- sorttbl[sorttbl$Petal.Length > 6.5,]
head(newtbl)

```

```{r}
newtbl2 <- newtbl[newtbl$Petal.Width < 2.2,]
head(newtbl2)

```

```{r}
newtbl3 <- sorttbl[(sorttbl$Petal.Length > 6.5) & (sorttbl$Petal.Width < 2.2),]
head(newtbl3)

```

```{r}
attributes(newtbl3)
```

```{r}
tbl[(tbl$Petal.Length + tbl$Petal.Width) * 2 > 17.5,]
```

```{r}
tbl <- defCasTable(conn, 'iris')
virginica <- tbl[tbl$Species == 'virginica',]
dim(virginica)

```

```{r}
head(virginica)
```
# Data Wrangling on the Fly

## Creating Computed Columns
```{r}
sorttbl['Sepal.Factor'] <- (sorttbl$Sepal.Length+sorttbl$Sepal.Width)*2
head(sorttbl)

```

```{r}
sorttbl['Total.Factor'] <- sorttbl$Sepal.Factor + sorttbl$Petal.Width + sorttbl$Petal.Length
head(sorttbl)

```

```{r}
sorttbl['names'] <- 'sepal / petal'
head(sorttbl)

```

```{r}
tbl = defCasTable(conn, 'iris')
tbl$is.virginica  <- (tbl$Species == 'virginica')
head(tbl) 

```
## BY-Group Processing
```{r}
tbl = defCasTable(conn, 'iris')
```

```{r}
grptbl <- tbl
grptbl@groupby <- 'Species'
attributes(grptbl)

```

```{r}
cas.simple.summary(grptbl, subset = c("min","max"))
```

```{r}
cas.simple.summary(
  conn, 
  table = list(name = 'iris', groupby = 'Species'),
  subset = c("min","max"))

```

### Concatenating BY Groups
```{r}
grpsumm <- cas.simple.summary(grptbl, subset = c("min","max"))
```

```{r}
rbind.bygroups(grpsumm)
```

Selecting Result Keys by Table Name
```{r}
grpcorr <- cas.simple.correlation(grptbl)
grpcorr

```

```{r}
result <- rbind.bygroups(grpcorr)
result$Correlation

```
### Handling Multiple Sets of BY Groups
```{r}
grpmdsumm <- cas.simple.mdSummary(
  tbl, 
  sets=list(
    list(groupby='Sepal.Length'),
    list(groupby='Petal.Length')
    )
  )
names(grpmdsumm)

```
